Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2018-10-17 documentation/paper.tex
Mostly minor edits

\begin{enumerate}
    \item Determine the most effective machine learning architecture to use in development of a discriminator. Furthermore, develop several neural network architectures and apply them to different input formats, in order to find which are the most effective.

@@

In this study, all newly developed discriminators were compared to the two discrimination techniques used previously for the PICO-60 experiment: Acoustic Parameter and machine learning.

@@

\subsubsection{Machine Learning}

@@

Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original neural analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands. This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.

@@

A certain number of cuts are necessarily applied to all data to ensure meaningful results. Otherwise, significant overfitting on biases in the data is likely. This basic quality cut $QualCut(D)$ consists of the following restrictions:

@@

Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel. It can be desirable for a discriminator to handle these events correctly; however, AP does not, and neither does the neural network used in the previous PICO-60 paper. Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.

@@

(Note that step \ref{exponent} only works if $p$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$. To resolve this, a piecewise exponential function is used instead, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards. This permits use of the full range of curves produced by even and non-integer values of $\psi$.)

@@

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% 90\% & 0.78 \\
            \hline
            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
            \hline
            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
            \hline
            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Iterative cluster nucleation (ICN) is a second original semi-supervised learning algorithm developed as part of this study. It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.

>>>>

\begin{enumerate}
    \item Determine the most effective machine learning architecture to use in development of a discriminator. Furthermore, develop several neural network architectures and apply them to different input formats (such as audio Fourier transforms, raw waveforms, and images), in order to find which are the most effective.

@@

In this study, all newly developed machine learning-based discriminators were compared to the two discrimination techniques used previously for the PICO-60 experiment: Acoustic Parameter and machine learning.

@@

\subsubsection{Original Machine Learning Analysis}

@@

Beyond this, the banded Fourier transform $\beta_{8}$ (which is the input to the AP function and the original machine learning analysis) was computed by integrating the resonant energy over all frequencies within each of a set of bands. This can be thought of as a method of signal downscaling, compressing the information into a smaller number of data points.

@@

A number of cuts are necessarily applied to all data to ensure meaningful results. Otherwise, significant overfitting on biases in the data is likely. This basic quality cut $QualCut(D)$ consists of the following restrictions:

@@

Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel. It can be desirable for a discriminator to handle these events correctly; however, AP alone does not, and neither does the neural network used in the previous PICO-60 paper. Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.

@@

(Note that step \ref{exponent} only works if $\phi$ is an odd integer (because odd integer exponents preserve the sign), limiting the ability to fine-tune this function by changing $\psi$. To resolve this, a piecewise exponential function is used instead, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards. This permits use of the full range of curves produced by even and non-integer values of $\psi$.)

@@

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
            \hline
            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
            \hline
            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
            \hline
            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Iterative cluster nucleation (ICN) is a second semi-supervised learning algorithm developed as part of this study. It takes advantage of some amount of imperfectly labeled data, using it to classify the rest of an unlabeled training set while optimizing to produce an effective discriminator.
